---
title: "R FINAL PROJECT"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(leaflet)
library(DT)
library(plotly)
library(flexdashboard)
library(sf)
library(tigris)


#Goal: to assess the diversity ratio among students across all the Higher Education institutions in Colorado

#plan: pivot data wider -> join datasets -> convert into sf object -> good to map
## data manipulation
# read data and filter just for Colorado

diversity_school <- read.csv("data/diversity_school.csv")

co_schools_data <- diversity_school %>% 
  filter(state == "Colorado") 

#Count amount of schools in Colorado
# co_schools_data %>% 
#   select(name) %>%
#   distinct() %>%
#   count()


#pivot wider so that each ethnic group has it's own column, and there is only one University name in the names column
co_schools_data_wider <- pivot_wider(co_schools_data,
            names_from = "category",
            values_from = enrollment)

#top ten universities in CO (based on enrollment)

top10_co_schools_enrollment <- co_schools_data_wider %>%
  distinct() %>%
  head(n= 10)


#coordinates for the top 10 schools in Colorado
top10_co_schools_coordinates <- read.csv("data/co_top10uni_coordinates.csv")



#change school name column name to match in both datasets

# top10_co_schools_coordinates
# top10_co_schools_enrollment

top10_co_schools_coordinates
top10_co_schools_enrollment


top10_co_schools_enrollment <- top10_co_schools_enrollment %>%
  rename(school = name)
#ok, now both have school names listed in column "school". Now join.


#merge data sets
top10CoData <- left_join(top10_co_schools_coordinates, top10_co_schools_enrollment)

top10CoData
```

### DATA DESCRIPTION

#TidyTuesday: School Diversity 

GOAL of the data: To explore the changing makeup of the nation’s school districts
To study ethnic diversity in United States schools, The Post measured diversity and integration using student race data from two years: 1995 and 2017.

The Post looked at the racial makeup of each of the country’s public school districts in 1995 and 2017 to see how the growing diversity of children was reflected in their schools. Private and charter schools were excluded from this analysis.

This dataset explores the proportions of students attending school in three categories: Unidiverse, Diverse and Extremely Diverse per state.

Diversity is a measure of the overall racial makeup of a school district. A district is considered diverse when no one race constitutes more than 75 percent of the school system’s student body overall.

Integration determines how well that diversity is reflected across the individual schools. An integrated school leads to better academic performance for students of color without affecting white students’ performance, according to decades of research analyzing test scores and graduation rates. And it offers an array of cultural and social benefits to all students, regardless of race.



GOAL of our flex dashboard

### INTERACTIVE MAP OF COLORADO SCHOOLS


```{r fig.height=8, fig.width=12, options(tigris_use_cache = TRUE)}
top10CoData <-top10CoData %>%
  mutate(longitude = top10CoData$longitude*(-1))

content <- paste(sep = "<br/>",
 top10CoData$school,
 paste("Total Enrollment = ", top10CoData$total_enrollment),
 paste("Total Minority = ", top10CoData$`Total Minority`))


leaflet() %>%
addTiles() %>%
fitBounds(-102.03,37,-109.03, 41) %>%
addTiles() %>%
addMarkers(top10CoData$longitude,
top10CoData$latitude,
popup = paste(sep = " ",
                  label = content))
```



### MOSAIC PLOT: ENROLLMENT DIVERSITY

```{r fig.asp=.5, fig.width=22, fig.align = "center"}        
#install.packages("ggmosaic")
library(ggplot2)
library(ggmosaic)


#pivot longer to obtain enrollment counts from "background"
top10formosiac <- top10_co_schools_enrollment %>%
  pivot_longer(cols = Women:`Non-Resident Foreign`,
               names_to = "Background",
               values_to = "enrollment")
#building a mosaic plot
mosaic <- top10formosiac %>%
  ggplot() +
  geom_mosaic(aes(x=product(school), fill = Background, na.rm = TRUE, 
                  weight = enrollment), offset = 0.008) + 
  theme_minimal() +
#aesthetic elements  
  theme(plot.title=element_text(size=22), legend.position="bottom", 
        legend.key.size = unit(0.2, 'cm'),legend.title = element_text(size=16), 
        legend.text = element_text(size=12), aspect.ratio = .4, 
        axis.text.x = element_text(angle = 45, hjust = 1, size = 16), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
#reversing legend order and adding titles
  scale_fill_discrete(guide = guide_legend(reverse=TRUE)) +
  labs(y= " ", x= " ", 
       title = "Visualizing Diversity in Colorado: 10 Largest Universities") 
#adding enrollment counts to mosaic squares
mosaic <- mosaic + geom_text(data = ggplot_build(mosaic)$data[[1]],
            aes(x = (xmin+xmax)/2, y = (ymin+ymax)/2, label=.wt))
mosaic

```

### SALARY OUTCOME 

```{r fig.height=10, fig.width=12}

#Reading in all potential data sets
diversity_school <- read.csv("data/diversity_school.csv")
all_schools <- read.csv("data/all-schools.csv")
historical_tuition <- read.csv("data/historical_tuition.csv")
salary_potential  <- read.csv("data/salary_potential.csv")
student_diversity <- read.csv("data/student_diversity.csv")
tuition_cost <- read.csv("data/tuition_cost.csv")
tuition_income <- read.csv("data/tuition_income.csv")

#Filtering down to just Colorado Schools 
library(dplyr)

co_school_salary_data <- salary_potential %>% 
  filter(state_name == "Colorado") 

#co_school_salary_data

#Top 10 Colorado Schools with Highest Early Career Pay 
top10_co_early_salary <-  co_school_salary_data %>%
  arrange(desc(early_career_pay)) %>%
  head(n=10)

top10_co_early_salary %>%
  datatable()

```


